---
title: "Jorge_MLBrainstem"
author: "Jorge"
date: "2023-06-23"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(dplyr)
```
```{r}
library(readxl)
xl_NewDenC <- read_excel("xl_NewDenC.xlsx")
```
```{r}
  xl_NewIntC <- subset(xl_NewIntC, select = -c(...1, voxRowNum, X, Y, Z))
```

```{r}
#filtering through data
TEST_143 <- filter(xl_NewDenC,`Structure-ID` == 143)

#Removing Structure-ID label
TEST_143 <- TEST_143 %>%
  select(-c(`Structure-ID`))

#Size of data
dim(TEST_143)
```

```{r}
#New data frame created
head(TEST_143)
```

##### **Generating Correlation Matrix**

Below is the code responsible for generating the correlation matrix of Structure-ID: 143.

```{r}
library(corrplot)
```

```{r}
library(doMC) 
```

```{r}
#Creating correlation matrix
correlation_143 <- cor(TEST_143)


#Adding logical vector that indicates which genes are N/A
is_na <- is.na(correlation_143)

#Pulling the names of the columns in the correlation matrix
column_names <- rownames(correlation_143)

#Finding the indices of the logical vector that are TRUE
indices_Prune <- which(is_na)

#Pulling the names of the indices that were found
na_columns <- column_names[indices_Prune]

#Printing the names of the columns that have N/A
head(na_columns, 75) 
```

```{r}
#Removing Inactive Genes
nu_TEST_143 <- TEST_143 %>%   select(-c(`Cnih3-39`, `Adcy9-83`, `Ap2b1-96`, `Pofut2-147`, `Rbbp6-170`, `Col6a1-290`, `Dbh-326`, `Erdr1-378`, `Enc1-390`, `Epha7-402`, `Extl3-411`, `Gata2-485`, `Hars-554`, `Hlf-565`, `Jam2-602`, `L1cam-620`, `Lhx1-633`, `Lig1-637`, `Marcksl1-670`, `Odz3-752`, `Odz4-754`, `Pcsk1n-777`, `Pkia-805`, `Plxnc1-820`, `Prkcz-841`, `Rbm4-888`, `Sema4a-942`, `Sst-1001`, `Vamp2-1092`, `Vipr2-1104`, `Ywhag-1124`, `Ctsl-1187`, `Cyr61-1191`, `Nr1d2-1245`, `Morc3-1311`, `4932438H23Rik-1321`, `Pex14-1397`, `Tmem2-1409`, `Etv3-1494`, `Pla2g12a-1518`, `Ube2g2-1547`, `Wdr18-1551`, `Plxnb1-1561`, `Casp9-1681`, `Trip4-1924`, `Snhg7-1968`, `6330407J23Rik-1988`, `Rfx7-1998`, `Ankrd33-2006`, `Adrm1-2019`, `Crat-2119`, `Csrp1-2129`, `Elk1-2185`, `Fbln1-2205`, `Hax1-2303`, `Irx1-2325`, `Ndufs1-2421`, `Nfkbie-2425`, `Park2-2462`, `Traf7-2527`, `Sla-2569`, `Aurka-2643`, `Mapk3-32576`, `Pim2-32583`, `Arxes1-45841`, `Limk2-61002`, `Bcap29-104826`, `Mlx-227539`, `Syn1-227540`, `Hhip-71358556`, `Itga7-73520991`, `Dbnl-74819497`, `Ggt7-75145998`, `Thbs4-77874596`, `Pcp2-79490073`))
```

```{r}
#Creating correlation matrix again
nu_correlation_143 <- cor(nu_TEST_143)
```

#### **Screening Through Correlation Matrix**

##### **Correlation Threshold: 90%**

Below is my attempt at screening through the produced correlation matrix (nu_correlation_143) and returning specific gene pairs whose correlation coefficient values are equal to or greater than 90%.

```{r}
library(caret)
```

```{r}
# METHOD 1 #
#Screening through matrix and returning columns (genes) that have 90% correlation or greater 
highlyCorrelated90 <- findCorrelation(nu_correlation_143, cutoff = .90, names = TRUE)
    
#Printing a sample of the returned columns (genes)
head(highlyCorrelated90)
```

```{r}
tail(highlyCorrelated90)
```

```{r}
# METHOD 2 #
#Finding all column pairs that have a correlation coefficient greater than 0.9
correlated_pairs90 <- which(abs(nu_correlation_143) > 0.9, arr.ind = TRUE)

#Printing the number of pairs found in total       
totalAT90 <- length(correlated_pairs90) - 4269
totalAT90
```

```{r}
#Printing a sample of the specified column pairs.
head(correlated_pairs90)
```

```{r}
tail(correlated_pairs90)
```

#### **Screening Through Correlation Matrix**

##### **Correlation Threshold: 80%**

Below is my attempt at screening through the produced correlation matrix (nu_correlation_143) and returning specific gene pairs whose correlation coefficient values are equal to or greater than 80%.

```{r}
# METHOD 1 
#Screening through matrix and returning columns (genes) that have 80% correlation or greater
highlyCorrelated80 <- findCorrelation(nu_correlation_143, cutoff = .80, names = TRUE)

#Printing a sample of the returned columns (genes)
head(highlyCorrelated80)
```

```{r}
tail(highlyCorrelated80)
```

```{r}
# METHOD 2 
#Finding all column pairs that have a correlation coefficient greater than 0.8
correlated_pairs80 <- which(abs(nu_correlation_143) > 0.8, arr.ind = TRUE)
  
#Printing the number of pairs found in total 
totalAT80 <- length(correlated_pairs80) - 4269
totalAT80
```

```{r}
#Printing a sample of the specified column pairs.
head(correlated_pairs80)
```

```{r}
tail(correlated_pairs80)
```

#### **Screening Through Correlation Matrix**

##### **Correlation Threshold: 70%**

Below is my attempt at screening through the produced correlation matrix (nu_correlation_143) and returning specific gene pairs whose correlation coefficient values are equal to or greater than 70%.

```{r}
# METHOD 1 
#Screening through matrix and returning columns (genes) that have 70% correlation or greater
highlyCorrelated70 <- findCorrelation(nu_correlation_143, cutoff = 0.70, names = TRUE)

#Printing a sample of the returned columns (genes)
head(highlyCorrelated70)
```

```{r}
tail(highlyCorrelated70)
```

```{r}
# METHOD 2 
#Finding all column pairs that have a correlation coefficient greater than 0.7
correlated_pairs70 <- which(abs(nu_correlation_143) > 0.7, arr.ind = TRUE)

#Printing the number of pairs found in total 
totalAT70 <- length(correlated_pairs70) - 4269
totalAT70
```

```{r}
#Printing a sample of the specified column pairs.
head(correlated_pairs70)
```

```{r}
tail(correlated_pairs70)
```
