---
title: "Jorge_MLBrainstem"
author: "Jorge"
date: "2023-06-23"
output:
  html_document: default
  pdf_document: default
---
```{r}
library(dplyr)
```
```{r}
library(readxl)
xl_NewDenC <- read_excel("xl_NewDenC.xlsx")
```
```{r}
 xl_NewDenC <- subset(xl_NewDenC, select = -c(...1, voxRowNum, X, Y, Z))
```

```{r}
#filtering through data
TEST_1098 <- filter(xl_NewDenC,`Structure-ID` == 1098)

#Removing Structure-ID label
TEST_1098 <- TEST_1098 %>%
  select(-c(`Structure-ID`))

#Size of data
dim(TEST_1098)
```

```{r}
#New data frame created
head(TEST_1098)
```

##### **Generating Correlation Matrix**

Below is the code responsible for generating the correlation matrix of Structure-ID: 1098.

```{r}
library(corrplot)
```

```{r}
library(doMC) 
```

```{r}
#Creating correlation matrix
correlation_1098 <- cor(TEST_1098)


#Adding logical vector that indicates which genes are N/A
is_na <- is.na(correlation_1098)

#Pulling the names of the columns in the correlation matrix
column_names <- rownames(correlation_1098)

#Finding the indices of the logical vector that are TRUE
indices_Prune <- which(is_na)

#Pulling the names of the indices that were found
na_columns <- column_names[indices_Prune]

#Printing the names of the columns that have N/A
head(na_columns, 144) 
```

```{r}
#Removing Inactive Genes
nu_TEST_1098 <- TEST_1098 %>%   select(-c(`Setd5-37`,`Cnih3-39`,`Adcy9-83`,`Arx-118`,`Avp-131`,`Camk2g-189`,`Cd1d1-219`,`Chrm4-261`,`Col6a1-290`,`Crh-292`,`Cxcl12-308`,`Dbh-326`,`Dlx3-350`,`Erdr1-378`,`Enc1-390`,`Entpd1-393`,`Fabp3-418`,`Grm1-537`,`Hars-554`,`Icmt-585`,`Kcnj6-609`,`L1cam-620`,`Lix1-643`,`Lpl-649`,`Neurod6-698`,`Nr3c1-727`,`Nrgn-736`,`Ntrk2-746`,`Ogt-758`,`Pcsk1n-777`,`Pkia-805`,`Prkcz-841`,`Slc6a3-978`,`Snca-986`,`Sst-1001`,`Syt7-1035`,`Th-1056`,`Uchl1-1088`,`Ywhag-1124`,`Ywhah-1127`,`Ywhaq-1130`,`Ywhaz-1133`,`Ikzf1-1144`,`Atp5o-1158`,`Cbfb-1168`,`Nr1d2-1245`,`ORF63-1247`,`Atm-1329`,`Cbs-1335`,`Elf2-1349`,`Esrra-1351`,`Lhx5-1369`,`Mx1-1385`,`Pex14-1397`,`Shh-1418`,`Gtf2h4-1458`,`Cpt1a-1478`,`Foxd3-1496`,`Lmx1b-1508`,`Mtf2-1510`,`Ndel1-1514`,`Suv39h2-1538`,`Wdr18-1551`,`Rwdd2b-1587`,`Homer1-1748`,`Tsen34-1944`,`Pttg1ip-1956`,`Snhg7-1968`,`6330407J23Rik-1988`,`Pde4dip-1994`,`Rfx7-1998`,`Ankrd33-2006`,`Adrm1-2019`,`Ankrd10-2027`,`Apbb1-2033`,`Agap3-2089`,`Cpt1c-2117`,`Crat-2119`,`Donson-2157`,`Erf-2195`,`Fbln1-2205`,`Gabarap-2245`,`Gnai2-2263`,`Gtf2i-2277`,`Irx1-2325`,`158303`,`Kpna4-2359`,`Leprotl1-2373`,`Lhx4-2377`,`Mapk13-2397`,`Mapre1-2401`,`Ndufs1-2421`,`Nfkbie-2425`,`Park2-2462`,`Pcsk1-2470`,`Pnrc1-2501`,`Ptn-2507`,`Traf7-2527`,`Leprel4-2547`,`Aurka-2643`,`Traf2-2691`,`Usp21-2711`,`Ltn1-2747`,`Mapk3-32576`,`Pim2-32583`,`Ube2i-45666`,`Dbndd2-45668`,`Arxes1-45841`,`Limk2-61002`,`Bcap29-104826`,`Vldlr-121089`,`Ptprcap-121091`,`Tnfrsf25-121096`,`Ssr2-121097`,`Gabpb1-121098`,`Pgf-121099`,`F2rl1-199391`,`Mef2a-199392`,`Tnfrsf11a-199393`,`Tpm3-275873`,`Atp6v0b-293682`,`Slco3a1-293691`,`Aqp3-293692`,`Zfand3-293694`,`Actr1a-321206`,`Pdcl-324213`,`Fbxw2-355882`,`Usp11-355883`,`Vps35-355884`,`Pls3-355886`,`Ncaph2-355887`,`Cct2-356082`,`Ptbp1-356857`,`Hectd1-357092`,`Trp53-67855414`,`BC100451-73615805`,`Cd8b1-73732149`,`Cyld-73992938`,`Uba6-74819486`,`Dbnl-74819497`,`Plcb1-75084969`,`Tmem91-75651161`,`Wnt5a-79556660`,`Cd6-112205643`))
```

```{r}
#Creating correlation matrix again
nu_correlation_1098 <- cor(nu_TEST_1098)
```

#### **Screening Through Correlation Matrix**

##### **Correlation Threshold: 90%**

Below is my attempt at screening through the produced correlation matrix (nu_correlation_1098) and returning specific gene pairs whose correlation coefficient values are equal to or greater than 90%.

```{r}
library(caret)
```

```{r}
# METHOD 1 #
#Screening through matrix and returning columns (genes) that have 90% correlation or greater 
highlyCorrelated90 <- findCorrelation(nu_correlation_1098, cutoff = .90, names = TRUE)
    
#Printing a sample of the returned columns (genes)
head(highlyCorrelated90)
```

```{r}
tail(highlyCorrelated90)
```

```{r}
# METHOD 2 #
#Finding all column pairs that have a correlation coefficient greater than 0.9
correlated_pairs90 <- which(abs(nu_correlation_1098) > 0.9, arr.ind = TRUE)

#Printing the number of pairs found in total       
totalAT90 <- length(correlated_pairs90) - 4200
totalAT90
```

```{r}
#Printing a sample of the specified column pairs.
head(correlated_pairs90)
```

```{r}
tail(correlated_pairs90)
```

#### **Screening Through Correlation Matrix**

##### **Correlation Threshold: 80%**

Below is my attempt at screening through the produced correlation matrix (nu_correlation_1098) and returning specific gene pairs whose correlation coefficient values are equal to or greater than 80%.

```{r}
# METHOD 1 
#Screening through matrix and returning columns (genes) that have 80% correlation or greater
highlyCorrelated80 <- findCorrelation(nu_correlation_1098, cutoff = .80, names = TRUE)

#Printing a sample of the returned columns (genes)
head(highlyCorrelated80)
```

```{r}
tail(highlyCorrelated80)
```

```{r}
# METHOD 2 
#Finding all column pairs that have a correlation coefficient greater than 0.8
correlated_pairs80 <- which(abs(nu_correlation_1098) > 0.8, arr.ind = TRUE)
  
#Printing the number of pairs found in total 
totalAT80 <- length(correlated_pairs80) - 4200
totalAT80
```

```{r}
#Printing a sample of the specified column pairs.
head(correlated_pairs80)
```

```{r}
tail(correlated_pairs80)
```

#### **Screening Through Correlation Matrix**

##### **Correlation Threshold: 70%**

Below is my attempt at screening through the produced correlation matrix (nu_correlation_1098) and returning specific gene pairs whose correlation coefficient values are equal to or greater than 70%.

```{r}
# METHOD 1 
#Screening through matrix and returning columns (genes) that have 70% correlation or greater
highlyCorrelated70 <- findCorrelation(nu_correlation_1098, cutoff = 0.70, names = TRUE)

#Printing a sample of the returned columns (genes)
head(highlyCorrelated70)
```

```{r}
tail(highlyCorrelated70)
```

```{r}
# METHOD 2 
#Finding all column pairs that have a correlation coefficient greater than 0.7
correlated_pairs70 <- which(abs(nu_correlation_1098) > 0.7, arr.ind = TRUE)

#Printing the number of pairs found in total 
totalAT70 <- length(correlated_pairs70) - 4200
totalAT70
```

```{r}
#Printing a sample of the specified column pairs.
head(correlated_pairs70)
```

```{r}
tail(correlated_pairs70)
```
