---
title: "Jorge_MLBrainstem"
author: "Jorge"
date: "2023-06-23"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(dplyr)
```
```{r}
library(readxl)
xl_NewDenC <- read_excel("xl_NewDenC.xlsx")
```
```{r}
 xl_NewDenC <- subset(xl_NewDenC, select = -c(...1, voxRowNum, X, Y, Z))
```

```{r}
#filtering through data
TEST_661 <- filter(xl_NewDenC,`Structure-ID` == 661)

#Removing Structure-ID label
TEST_661 <- TEST_661 %>%
  select(-c(`Structure-ID`))

#Size of data
dim(TEST_661)
```

```{r}
#New data frame created
head(TEST_661)
```

##### **Generating Correlation Matrix**

Below is the code responsible for generating the correlation matrix of Structure-ID: 661.

```{r}
library(corrplot)
```

```{r}
library(doMC) 
```

```{r}
#Creating correlation matrix
correlation_661 <- cor(TEST_661)


#Adding logical vector that indicates which genes are N/A
is_na <- is.na(correlation_661)

#Pulling the names of the columns in the correlation matrix
column_names <- rownames(correlation_661)

#Finding the indices of the logical vector that are TRUE
indices_Prune <- which(is_na)

#Pulling the names of the indices that were found
na_columns <- column_names[indices_Prune]

#Printing the names of the columns that have N/A
head(na_columns, 19) 
```

```{r}
#Removing Inactive Genes
nu_TEST_661 <- TEST_661 %>%   select(-c(`Col6a1-290`, `Extl3-411`, `Lig1-637`, `Nrtn-741`, `Prdx1-836`, `Prkcz-841`, `Sst-1001`, `Vamp2-1092`, `Ywhag-1124`, `Tmem2-1409`, `Plxnb1-1561`, `Adrm1-2019`, `Crat-2119`, `Fbln1-2205`, `Ndufs1-2421`, `Traf7-2527`, `Pim2-32583`, `Limk2-61002`, `Bcap29-104826`))
```

```{r}
#Creating correlation matrix again
nu_correlation_661 <- cor(nu_TEST_661)
```

#### **Screening Through Correlation Matrix**

##### **Correlation Threshold: 90%**

Below is my attempt at screening through the produced correlation matrix (nu_correlation_661) and returning specific gene pairs whose correlation coefficient values are equal to or greater than 90%.

```{r}
library(caret)
```

```{r}
# METHOD 1 #
#Screening through matrix and returning columns (genes) that have 90% correlation or greater 
highlyCorrelated90 <- findCorrelation(nu_correlation_661, cutoff = .90, names = TRUE)
    
#Printing a sample of the returned columns (genes)
head(highlyCorrelated90)
```

```{r}
tail(highlyCorrelated90)
```

```{r}
# METHOD 2 #
#Finding all column pairs that have a correlation coefficient greater than 0.9
correlated_pairs90 <- which(abs(nu_correlation_661) > 0.9, arr.ind = TRUE)

#Printing the number of pairs found in total       
totalAT90 <- length(correlated_pairs90) - 4325
totalAT90
```

```{r}
#Printing a sample of the specified column pairs.
head(correlated_pairs90)
```

```{r}
tail(correlated_pairs90)
```

#### **Screening Through Correlation Matrix**

##### **Correlation Threshold: 80%**

Below is my attempt at screening through the produced correlation matrix (nu_correlation_661) and returning specific gene pairs whose correlation coefficient values are equal to or greater than 80%.

```{r}
# METHOD 1 
#Screening through matrix and returning columns (genes) that have 80% correlation or greater
highlyCorrelated80 <- findCorrelation(nu_correlation_661, cutoff = .80, names = TRUE)

#Printing a sample of the returned columns (genes)
head(highlyCorrelated80)
```

```{r}
tail(highlyCorrelated80)
```

```{r}
# METHOD 2 
#Finding all column pairs that have a correlation coefficient greater than 0.8
correlated_pairs80 <- which(abs(nu_correlation_661) > 0.8, arr.ind = TRUE)
  
#Printing the number of pairs found in total 
totalAT80 <- length(correlated_pairs80) - 4325
totalAT80
```

```{r}
#Printing a sample of the specified column pairs.
head(correlated_pairs80)
```

```{r}
tail(correlated_pairs80)
```

#### **Screening Through Correlation Matrix**

##### **Correlation Threshold: 70%**

Below is my attempt at screening through the produced correlation matrix (nu_correlation_661) and returning specific gene pairs whose correlation coefficient values are equal to or greater than 70%.

```{r}
# METHOD 1 
#Screening through matrix and returning columns (genes) that have 70% correlation or greater
highlyCorrelated70 <- findCorrelation(nu_correlation_661, cutoff = 0.70, names = TRUE)

#Printing a sample of the returned columns (genes)
head(highlyCorrelated70)
```

```{r}
tail(highlyCorrelated70)
```

```{r}
# METHOD 2 
#Finding all column pairs that have a correlation coefficient greater than 0.7
correlated_pairs70 <- which(abs(nu_correlation_661) > 0.7, arr.ind = TRUE)

#Printing the number of pairs found in total 
totalAT70 <- length(correlated_pairs70) - 4325
totalAT70
```

```{r}
#Printing a sample of the specified column pairs.
head(correlated_pairs70)
```

```{r}
tail(correlated_pairs70)
```
